Introductory Physics Labs: A Tale of Two Transformations

### 2021-03-30 manuscript/onlineLabPaper.tex
Revisions in Word file and in LaTeX file.

\subsection{Session 1: Pre-Lab}
Part of the goal of any laboratory is to help students develop a familiarity with certain tools
and equipment common to the lab.  The purpose of the pre-lab reading and activity is to give
students familiarity with the equipment they will be using.  For the investigation described
previously, ``Does the force a fan exerts on a cart depend on the mass of the cart?''  students 
must become familiar with a few things, such as what we mean by a cart and also obtain some 
tools to be able to answer this question.  As a part of the pre-lab
reading, students learn about linear regression---both how to run it in Excel,
how to interpret the output, and how we can estimate instantaneous velocity given position
vs.\ time data. For this experiment, we give them a cart (leaving the fan off for now) and have
them measure the acceleration of that cart down a ramp using an ultrasonic motion detector to
collect position vs.\ time data.  Students work in pairs and submit a brief
summary of this work on the learning management system (LMS). This investigation is the first
time the students create graphs, so we set norms for graphing such as helping them decide which
variable to put on the $x$-axis and the $y$-axis as well as including units.  Pre-lab assignments
are designed to be graded in less than 1 minute/submission based on a rubric we provide to both
students and GTAs.

@@

\subsection{Session 3: Argumentation and Peer Review}
During this session, the students share their results in a poster session.  Students finish analyzing
their data during the prior week, and prepare a ``poster'' on a whiteboard.  One person from
the group stays at the whiteboard while the other group members (travelers) go to other groups.
The travelers learn what other groups did, and compare to their results in a structured
argumentation setting.  Groups spend about 3--5 minutes at a poster, and rotate so that they
see at least 3 other posters.  After the travelers return, critically, the groups spend time
discussing what they saw and reflecting on their own results.  What is exciting is that
sometimes, these interactions actually drive groups to change their claims.\citep{Walker2019}

@@

Proposal development also occurred online asynchronously.  Each group produced their proposal
and posted a proposal form on a discussion board for approval. The GTA reviewed proposals and
provided feedback or approval. Groups used the GTA feedback to revise their proposals until
they were approved. As with the prior face-to-face labs, the GTAs were looking for specific items in 
the proposals such as what data the students planned to collect, how they would use the data to 
answer the guiding question, what steps they would take to minimize errors, and how the students 
would quantify their uncertainty. Most proposals obtained approval after two or three revisions, but 
some required as many as seven revisions to obtain approval. When the GTA approved a group's
proposal, they assigned the group a data set for the investigation, and the group began
measurements and analysis. Groups were encouraged to revise their methods during the  
measurement/analysis phase if they discovered a flaw or a better way to answer the guiding question. 
Based on our prior experience running the class, we prepared some
data sets appropriate to the different proposals that previous students had produced and
provided measurements to the current students in the most raw form possible to require them to
make decisions about data collection and analysis. We provided six data sets for the third investigation 
and 13 data sets for the fourth investigation for the Physics 1 laboratory. We provided seven data sets 
for the third investigation (each group was assigned two data sets at random to compare) and 14 data 
sets for the fourth investigation for the Physics 2 laboratory.

@@

\section{Discussion}
We have observed a great number of successes for these transformations.  First and foremost, we
indeed have a lab curriculum that is aligned with national goals and standards. Furthermore, we
have successfully deployed this online.  Students are engaging with science practices as
determined by our practical assessments, and our department culture around laboratory course
instruction is invigorated. Psychometric assessment of our students engagement with science
practices is ongoing.

>>>>

\subsection{Session 1: Pre-Lab}
Part of the goal of any laboratory is to help students develop a familiarity with certain tools
and equipment common to the lab.  The purpose of the pre-lab reading and activity is to give
students familiarity with the equipment they will be using.  For the investigation described
previously, ``Does the force a fan exerts on a cart depend on the mass of the cart?''  students 
must become familiar with a few things, such as what we mean by a cart and also obtain some 
tools to be able to answer this question.  As a part of the pre-lab
reading, students learn about linear regression---how to run it in Excel,
how to interpret the output, and how we can estimate instantaneous velocity given position
vs.\ time data. For this experiment, we give them a cart (leaving the fan off for now) and have
them measure the acceleration of that cart down a ramp using an ultrasonic motion detector to
collect position vs.\ time data.  Students work in pairs and submit a brief
summary of this work on the learning management system (LMS). This investigation is the first
time the students create graphs, so we set norms for graphing such as helping them decide which
variable to put on the $x$-axis and the $y$-axis as well as including units.  Pre-lab assignments
are designed to be graded in less than 1 minute/submission based on a rubric we provide to both
students and GTAs.

@@

\subsection{Session 3: Argumentation and Peer Review}
During this session, the students share their results in a poster session.  Students finish analyzing
their data during the prior week and prepare a ``poster'' on a whiteboard.  One person from
the group stays at the whiteboard while the other group members (travelers) go to other groups.
The travelers learn what other groups did and compare to their results in a structured
argumentation setting.  Groups spend about 3--5 minutes at a poster and rotate so that they
see at least 3 other posters.  After the travelers return, critically, the groups spend time
discussing what they saw and reflecting on their own results.  What is exciting is that
sometimes, these interactions actually drive groups to change their claims.\citep{Walker2019}

@@

Proposal development also occurred online asynchronously.  Each group produced their proposal
and posted a proposal form on a discussion board for approval. The GTA reviewed proposals and
provided feedback or approval. Groups used the GTA feedback to revise their proposals until
they were approved. As with the prior face-to-face labs, the GTAs were looking for specific items in 
the proposals such as what data the students planned to collect, how they would use the data to 
answer the guiding question, what steps they would take to minimize errors, and how the students 
would quantify their uncertainty. Most proposals obtained approval after two or three revisions, but 
some required as many as seven revisions to obtain approval. When the GTA approved a group's
proposal, they assigned the group a data set for the investigation and the group began
measurements and analysis. Groups were encouraged to revise their methods during the  
measurement/analysis phase if they discovered a flaw or a better way to answer the guiding question. 
Based on our prior experience running the class, we prepared some
data sets appropriate to the different proposals that previous students had produced and
provided measurements to the current students in the most raw form possible to require them to
make decisions about data collection and analysis. We provided six data sets for the third investigation 
and 13 data sets for the fourth investigation for the Physics 1 laboratory. We provided seven data sets 
for the third investigation (each group was assigned two data sets at random to compare) and 14 data 
sets for the fourth investigation for the Physics 2 laboratory.

@@

\section{Discussion}
We have observed a great number of successes for these transformations.  First and foremost, we
indeed have a lab curriculum that is aligned with national goals and standards. Furthermore, we
have successfully deployed this online.  Students are engaging with science practices as
determined by our practical assessments, and our department culture around laboratory course
instruction is invigorated. Psychometric assessment of our students' engagement with science
practices is ongoing.
